The role of advection versus diffusion in intraventricular transport

### 2025-02-04 paper/main.tex
Results 1

\mer{MER: @Halvor I suggest you move these two paragraphs into their respective Results sections (Results 2 and 5). Putting the bullet points above at the end of the Stokes equations is a great idea!}
When simulating transport of solute in the ventricles, we consider the same six different regions of interest that were considered in the photoconversion experiments. In region number one (the photoconversion region), we use the function~\eqref{eq:photoconversion_curve} to enforce the concentration such that it represents the experimental data of photoconversion of proteins. We consider the three different diffusion coefficients introduced in previous sections to simulate and compare results for transport of molecules of varying sizes.

@@

To evaluate the relative importance of advection and diffusion as transport mechanisms, we consider the Péclet number
\begin{equation*}
    \mathrm{Pe} = \frac{\mathrm{diffusion \ timescale}}{\mathrm{advection \ timescale}} = \frac{t_d}{t_a} = \frac{L_p^2/D}{L_p/U_p} = \frac{L_p U_p}{D}.
\end{equation*}
Here, $L_p$ and $U_p$ are characteristic length and velocity scales, respectively, and $D$ is the solute diffusion coefficient. \fixme{Thus, if $\mathrm{Pe} > 1$, transport is dominated by advection, and if $\mathrm{Pe} < 1$ transport is dominated by diffusion. As a length scale $L_p$, we choose the approximate length of the ventricles geometry in the $x$-direction, $L_p=600$ \textmu m. We use a mean velocity $\overline{U}$ as the characteristic velocity $U_p$. The mean velocity $\overline{U}$ is calculated by first averaging the velocity in space:
\begin{equation*}
    \overline{u}^2 = \frac{1}{|\Omega|}\int_{\Omega}\uu\cdot\uu\dx,
\end{equation*}
and then averaging $\overline{u}$ in time over one cardiac cycle.}

To assess the assumption that the CSF flow is governed by the Stokes equations, we also compute the Reynolds number
\begin{equation*}
    \mathrm{Re} = \frac{\mathrm{inertial \ forces}}{\mathrm{viscous \ forces}} = \frac{\rho U_r L_r}{\mu},
\end{equation*}
with $L_r$ and $U_r$ as length and velocity scales characteristic of the flow.

@@

The Stokes equations~\eqref{eq:stokes_eqs} are discretized with a finite element method in space, while in time, we employ a first-order implicit Euler discretization.
We set the timestep size $\Delta t = 0.02252$ s such that 20 timesteps represent one cardiac cycle. For the spatial discretization, we use piecewise linear Brezzi-Douglas-Marini (BDM) elements for the velocity.
The BDM elements have degrees of freedom associated with integral moments of the velocity component normal to the facets of elements~\cite{Brezzi1985TwoProblems}, thus we enforce the impermeability condition $\uu\cdot\nn=0$ strongly.
We impose the boundary conditions for both the tangential traction $\bsigpar$ and the normal pressure $\Tilde{p}(t)$ weakly.
For the pressure, we use piecewise constant (zeroth order discontinuous Galerkin (DG)) elements.
The resulting total number of degrees of freedom at each time step is 943 820.
\fixme{The element combination $\mathrm{BDM}_1$--$\mathrm{DG}_{0}$ yields a non-conforming but stable discretization scheme for the Stokes equations~\cite{Stenberg1989SomeEquations}.
The discretization includes a penalty parameter $\gamma$ which needs to be chosen large enough to ensure stability of the finite element method. We set $\gamma=10$.}
Since the divergence of the velocity is a subspace of the pressure space for this scheme, the divergence-free condition of the velocity is satisfied exactly on each element~\cite{Boffi2008FiniteProblem}, ensuring mass conservation.

@@

\subsection*{Solution strategy and software implementation}
Because of the one-way coupling between the velocity field $\uu$ and the concentration $c$, the governing equations for the CSF flow and the concentration can be solved sequentially when modeling solute transport.
\fixme{For the cilia-only/no-cardiac flow scenario, the velocity quickly reaches steady-state.} In this case, we first solve the \fixme{steady-state} Stokes equations~\eqref{eq:stokes_eqs} for $\uu$ and $p$,
and then use the velocity field $\uu$ as input when solving the advection-diffusion equation~\eqref{eq:adv_diff_strong}.
For the other flow scenarios, the Stokes equations are solved for one period of the pulsatile cardiac motion. The solution of this one period is then used periodically as input for the advection-diffusion equation.

@@

The larval and adult zebrafish were raised under standard husbandry conditions at 28.5°C in a Techniplast Zebtech Multilinking system.
The fish tanks were kept at constant pH 7.0 and 685 $\mu$S with a 14/10 hr light/dark cycle.
From fertilization to 3 dpf, larvae were maintained in egg water (1.2 g marine salt and 0.1\% methylene blue in 20 L reverse osmosis (RO) water)
and subsequently transferred to artificial fish water (AFW) (1.2 g marine salt in 20 L RO water).
The zebrafish lines used in the experiments were \emph{ccdc103(dnaaf19)\textsuperscript{tn222a}} (\emph{schmalhans (smh)})~\cite{Jau-NianChen1997Left-rightZebrafish}
and \emph{Tg(gfap:Gal4FF)\textsuperscript{nw7Tg}}~\cite{DiazVerdugo2019Glia-neuronSeizures}, \emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}}.
Animals were in the pigmentless \emph{nacre\textsuperscript{b692}}(\emph{mitfa\textsuperscript{-/-}})~\cite{JamesA.Lister1999NacreFate} background.

@@

%%--------------RESULTS--------------%%
\section*{Results}\label{sec:results}

@@

\mer{One sentence on why? - what do we not know - what do we try to
  find out?} \mer{@Halvor: If you want, try to revamp this paragraph
  as suggested by Nathalie. First cilia. Then pulsatile. Then
  combined.} To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we begin by
considering the simulated CSF velocity and pressure when driven by
both cilia motion and cardiac pressure pulsations. These forces induce
\fixme{persistent} flow features, with vortex-structures in the
anterior and middle ventricles and \fixme{unidirectional} flow in the
posterior ventricle (\Cref{fig:figure1}e,
\fixme{Supplementary video 1}). These characteristic patterns suggest
that the flow in the anterior and middle ventricles are dominated by
the cilia contribution, while the cardiac pressure pulsations dominate
within the posterior ventricle. Over one cardiac cycle, the CSF flow
speed reaches up to 27.7 \textmu m/s, dorsally in the posterior part
of the middle ventricle. This peak is attained at the time instant
coinciding with the time of peak caudal cardiac pulsatile flow, as a
result of both the traction induced by the cilia and the pulsatile
pressure-driven flow being aligned with the rostrocaudal ($x-$)
axis. Using this maximum velocity magnitude and choosing the height of
the middle ventricle ($110$ \textmu m) as a characteristic length
scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$,
justifying the assumption of Stokes flow. To further examine the
effect of the cilia, we next consider a no-cilia model variation. In
this scenario, the simulated CSF flow is pulsatile and
\fixme{unidirectional} (\Cref{fig:figure1}g), with a peak velocity
magnitude of \fixme{$4.8$ \textmu m/s}. On the other hand, the
vortices (re)appear when considering a variation with cilia but no
cardiac contribution (\Cref{fig:figure1}h), in which case the highest
velocities ($26.5$ \textmu m/s) also occur in the vicinity of the
cilia. Thus, the bidirectional traction induced by motile cilia
inherently leads to a partial compartmentalization of the ventricular
system.

%% %\subsection*{Vortex structures vanish in the absence of cilia}
%% The pulsatile motion is absent when disregarding cardiac motion~(model I, \cref{fig:figure1}, panel~\textbf{f}). For model I, the boundary conditions are not time-dependent, rendering the solution to the Stokes equations~\eqref{eq:stokes_eqs} a steady-state flow field. Streamlines demonstrate the compartmentalization of the cerebrospinal fluid flow driven by cilia, and we observe highest vorticity in the vicinity of cilia.

%% Excluding the cilia tangential traction from the model (model II), vortex-structured compartmentalization of the ventricular flow does not occur, and the flow is directional and pulsatile as a result of the pressure boundary conditions imposed~(\cref{fig:figure1}, panel \textbf{g}).

%% The cardiac pressure forces induce a pulsatile motion of the
%% cerebrospinal fluid synchronized with the cardiac cycle, with velocity
%% magnitudes of the pulsatile motion being approximately an order of
%% magnitude smaller than the velocities related to the cilia-driven
%% motion.

@@

\subsection*{Absence of ciliary motion affects local protein distribution}
To assess the impact of cilia motion on solute distribution, we measured the movement of photoconverted Dendra2 signal
in control and motile-cilia mutant zebrafish (\emph{schmalhans} (\emph{smh})).
By plotting the means of $\Delta F$ (\Cref{fig:figure4_compare_control_mutant}a) and the time to threshold
across the control and mutant cohorts (\Cref{fig:figure4_compare_control_mutant}b), we identified a significant
difference in Dendra2 distribution in ROIs 2 and 4, but not in ROIs 3, 5, and 6. In ROIs 2 and 4,
which measure Dendra2 distribution downstream of the photoconversion site,
we observed a slower increase of fluorescence intensity,
resulting in a $\sim 45\%$ higher mean time-to-threshold in the \emph{smh}
mutant compared to controls (\Cref{fig:figure4_compare_control_mutant}b).

Next, to identify whether our model would recapitulate these findings,
we simulated protein transport by using the only-cardiac/no-cilia flow model 
and compared it with the baseline model (both cilia and cardiac motion) (\Cref{fig:figure4_compare_control_mutant}c).
Similar to our experiments, we observed slower dynamics in simulations for ROIs 2 and 4.
Interestingly, the difference was even greater for the simulations,
with the time-to-threshold in ROI 2 being 90\% higher in absence of cilia-mediated flow.
Moreover, ROI 5 exhibits very different $\cbar$ profiles for the two configurations in our computational model,
but not in our experimental data.
Altogether, our findings identify that diffusion plays an important role for solute movement to regions upstream in the cilia flow fields or to distant ventricles.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F$ over time in each region of interest (ROI) for control (black lines) and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F$ first exceeds a threshold value, equal to 0.25 (ROIs 2-4) or 0.10 (ROIs 5, 6), for control (gray dots) and \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets, and $p$-values were calculated using a Mann-Whitney U test with a confidence level of 95\%.
    \textbf{c.} Mean concentration $\cbar$ as function of time in each ROI, simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have different limits in each plot.}
    \label{fig:figure4_compare_control_mutant}
\end{figure}

\subsection*{Loss of cilia motility in specific subpopulation slows down transport}
We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the role of these lineages on CSF transport, we modified flow boundary conditions
in the three different ciliated regions (\Cref{fig:figure5_compare_cilia_modifications}a) and simulated
CSF flow using the Stokes equations. We identified that the velocities were most strongly
affected when removing cilia in the dorsal diencephalic cilia,
as the maximum velocity dropped from 27.7 $\mu$m/s to 8.0 $\mu$m/s.
In contrast, removing telencephalic and ventral diencephalic had minimal
impact and resulted in maximum velocities of 27.7 $\mu$m/s or 27.2 $\mu$m/s, respectively.

With the resulting velocity fields, we simulated transport of Dendra2.
Changes in the transport dynamics were most prominent when removing the dorsal cilia
in the diencephalic ventricle, which produces the strongest flow at this developmental stage.
All ROIs (except ROI 1, where we locally released solutes) exhibit significant changes
in the dynamics of $\cbar$ (\Cref{fig:figure5_compare_cilia_modifications}b).
Compared to the baseline model results, we observed a slower increase and lower final concentration
in the middle and anterior ventricles (ROIs 2--5),
and a faster transport to the posterior ventricle (ROI 6).
In contrast, removal of telencephalic cilia did not alter the evolution of the
mean concentration $\cbar$ with time (\Cref{fig:figure5_compare_cilia_modifications}b).
Finally, without ventral cilia in the diencephalic ventricle,
more solute has spread towards the anterior ventricle (ROI 5) and
less solute has spread towards the posterior ventricle (ROI 6).

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{Caption}
    \label{fig:figure5_compare_cilia_modifications}
\end{figure}

>>>>

\mer{MER: @Halvor I suggest you move these two paragraphs into their respective
Results sections (Results 2 and 5).}
When simulating transport of solute in the ventricles,
we consider the same six different regions of interest that were
considered in the photoconversion experiments.
In region number one (the photoconversion region),
we use the function~\eqref{eq:photoconversion_curve} to enforce
the concentration such that it represents the experimental data
of photoconversion of proteins. We consider the three different diffusion
coefficients introduced in previous sections to simulate and compare results
for transport of molecules of varying sizes.

@@

To evaluate the relative importance of advection and
diffusion as transport mechanisms, we consider the Péclet number
\begin{equation*}
    \mathrm{Pe} = \frac{\mathrm{diffusion \ timescale}}{\mathrm{advection \ timescale}}
    = \frac{t_d}{t_a} = \frac{L_p^2/D}{L_p/U_p} = \frac{L_p U_p}{D}.
\end{equation*}
Thus, if $\mathrm{Pe} > 1$, transport is dominated by advection,
and if $\mathrm{Pe} < 1$ transport is dominated by diffusion.
Here, $L_p$ and $U_p$ are characteristic length and velocity scales,
respectively, and $D$ is the solute diffusion coefficient.
In calculating a global Péclet number, we use $L_p=600$ \textmu m, the approximate length of the
ventricles geometry along the rostrocaudal ($x$-) axis.
We use a mean velocity $\overline{U}$ as the characteristic velocity $U_p$.
The mean velocity $\overline{U}$ is calculated by first averaging the velocity in space:
\begin{equation*}
    \overline{u}^2 = \frac{1}{|\Omega|}\int_{\Omega}\uu\cdot\uu\dx,
\end{equation*}
and then averaging $\overline{u}$ in time over one cardiac cycle. To assess the assumption that the CSF flow is governed by the Stokes equations,
we also compute the Reynolds number
\begin{equation*}
    \mathrm{Re} = \frac{\mathrm{inertial \ forces}}{\mathrm{viscous \ forces}} = \frac{\rho U_r L_r}{\mu},
\end{equation*}
with $L_r$ and $U_r$ as length and velocity scales characteristic of the flow.

@@

The Stokes equations~\eqref{eq:stokes_eqs} are discretized with a finite element method
in space, while in time, we employ a first-order implicit Euler discretization.
We set the timestep size $\Delta t = 0.02252$ s such that 20 timesteps represent one cardiac cycle.
For the spatial discretization, we use piecewise linear Brezzi-Douglas-Marini (BDM) elements for the velocity.
The BDM elements have degrees of freedom associated with integral moments of the
velocity component normal to the facets of elements~\cite{Brezzi1985TwoProblems},
thus we enforce the impermeability condition $\uu\cdot\nn=0$ strongly.
We impose the boundary conditions for both the tangential traction $\bsigpar$ and
the normal pressure $\Tilde{p}(t)$ weakly.
For the pressure, we use piecewise constant (zeroth order discontinuous Galerkin (DG)) elements.
The resulting total number of degrees of freedom at each time step is 943 820.
\fixme{The element combination $\mathrm{BDM}_1$--$\mathrm{DG}_{0}$ yields a non-conforming
discretization scheme for the Stokes equations~\cite{Stenberg1989SomeEquations}.
To ensure stability of the scheme, continuity in tangential velocities on interior facets
is weakly enforced with a penalty parameter $\gamma=10$~\cite{Hong2016AEquations}.}
Since the divergence of the velocity is a subspace of the pressure space for this scheme,
the divergence-free condition of the velocity is satisfied exactly on each
element~\cite{Boffi2008FiniteProblem}, ensuring mass conservation.

@@

\subsection*{Solution strategy and software implementation}
Owing to the one-way coupling between the velocity field $\uu$
and the concentration $c$, the governing equations for the
CSF flow and the concentration can be solved sequentially when modeling solute transport.
\fixme{For the cilia-only/no-cardiac flow scenario, the velocity quickly reaches steady-state.}
In this case, we first solve the \fixme{steady-state} Stokes equations~\eqref{eq:stokes_eqs}
for $\uu$ and $p$, and then use the velocity field $\uu$ as input when solving
the advection-diffusion equation~\eqref{eq:adv_diff_strong}.
For the other flow scenarios, the Stokes equations are solved for one
period of the pulsatile cardiac motion. The solution of this one period
is then used periodically as input for the advection-diffusion equation.

@@

The larval and adult zebrafish were raised under standard husbandry
conditions at 28.5°C in a Techniplast Zebtech Multilinking system.
The fish tanks were kept at constant pH 7.0 and 685 $\mu$S with a
14/10 hr light/dark cycle.
From fertilization to 3 dpf, larvae were maintained in egg
water (1.2 g marine salt and 0.1\% methylene blue in 20 L reverse osmosis (RO) water)
and subsequently transferred to artificial fish water (AFW)
(1.2 g marine salt in 20 L RO water). The zebrafish lines used in the experiments were
\emph{ccdc103(dnaaf19)\textsuperscript{tn222a}}
(\emph{schmalhans (smh)})~\cite{Jau-NianChen1997Left-rightZebrafish}
and \emph{Tg(gfap:Gal4FF)\textsuperscript{nw7Tg}}~\cite{DiazVerdugo2019Glia-neuronSeizures},
\emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}}.
Animals were in the pigmentless \emph{nacre\textsuperscript{b692}}
(\emph{mitfa\textsuperscript{-/-}})~\cite{JamesA.Lister1999NacreFate} background.

@@

%%--------------RESULTS--------------%%
\section*{Results}

@@

We seek to model the compartimentalized and pulsatile CSF flow observed \emph{in vivo}~(\Cref{fig:figure1}b--c).
To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we begin by
considering the simulated CSF velocity and pressure when only considering
cilia motion, disregarding cardiac pressure pulsations. The
bidirectional traction induced by the motile cilia
inherently leads to a partial compartmentalization of the ventricular
system, with large-scale vortex structures in the anterior and middle
ventricles~(\Cref{fig:figure1}e). The highest velocities ($26.5$ \textmu m/s)
occur in the vicinity of the dorsal cilia in the middle ventricle.

To examine the effect of cardiac pressure pulsations,
we next consider a no-cilia model variation. In this scenario, the simulated CSF
flow is pulsatile as a result of the sinusoidal pressure gradient (\Cref{fig:figure1}g),
with a peak velocity magnitude of $4.8$ \textmu m/s. At each time instant, the flow
is unidirectional, with CSF initially flowing rostrocaudally (positive $x-$axis) before changing
direction to caudorostral in the middle of the cardiac cycle.

We now consider the baseline model, simulating flow driven by both cilia
and cardiac pulsations. We observe \fixme{persistent} flow features,
with vortex-structures in the anterior and middle ventricles
and directional flow in the posterior ventricle (\Cref{fig:figure1}e,
\fixme{Supplementary video 1}). These characteristic patterns suggest
that the flow in the anterior and middle ventricles are dominated by
the cilia contribution, while the cardiac pressure pulsations dominate
within the posterior ventricle. Over one cardiac cycle, the CSF flow
speed reaches up to 27.7 \textmu m/s, dorsally in the posterior part
of the middle ventricle. This peak is attained at the time instant
coinciding with the time of peak caudal cardiac pulsatile flow, as a
result of both the traction induced by the cilia and the pulsatile
pressure-driven flow being aligned with the rostrocaudal 
axis. Using the observed maximum velocity magnitude and choosing the height of
the middle ventricle ($110$ \textmu m) as a characteristic length
scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$,
justifying the assumption of Stokes flow. In summary, the computational
CSF flow model reproduces the flow features observed \emph{in vivo}.

@@

\subsection*{Absence of ciliary motion affects local protein distribution}
To assess the impact of cilia motion on solute distribution,
we measured the movement of photoconverted Dendra2 signal
in control and motile-cilia mutant zebrafish (\emph{schmalhans} (\emph{smh})).
By plotting the means of $\Delta F$ (\Cref{fig:figure4_compare_control_mutant}a)
and the time to threshold
across the control and mutant cohorts (\Cref{fig:figure4_compare_control_mutant}b),
we identified a significant
difference in Dendra2 distribution in ROIs 2 and 4, but not in ROIs 3, 5, and 6. In ROIs 2 and 4,
which measure Dendra2 distribution downstream of the photoconversion site,
we observed a slower increase of fluorescence intensity,
resulting in $\sim X\%$ (ROI 2) and $\sim 45\%$ (ROI 4)
higher mean times-to-threshold in the \emph{smh}
mutants compared to controls (\Cref{fig:figure4_compare_control_mutant}b).

Next, to identify whether our model would recapitulate these findings,
we simulated protein transport by using the only-cardiac/no-cilia flow model 
and compared it with the baseline model
(both cilia and cardiac motion) (\Cref{fig:figure4_compare_control_mutant}c).
Similar to our experiments, we observed slower dynamics in simulations for ROIs 2 and 4.
Interestingly, the difference was even greater for the simulations,
with the time-to-threshold in ROI 2 being 90\% higher in absence of cilia-mediated flow.
Moreover, ROI 5 exhibits very different $\cbar$ profiles for the two
configurations in our computational model,
but not in our experimental data.
Altogether, our findings identify that diffusion plays an important
role for solute movement to regions upstream in the cilia flow fields or to distant ventricles.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F$ over
    time in each region of interest (ROI) for control (black lines) and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets, and $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%.
    \textbf{c.} Mean concentration $\cbar$ as function of time in each ROI,
    simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:figure4_compare_control_mutant}
\end{figure}

\subsection*{Loss of cilia motility in specific subpopulation slows down transport}
We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the role of these lineages on CSF transport, we modified flow boundary conditions
in the three different ciliated regions
(\Cref{fig:figure5_compare_cilia_modifications}a, magenta markers)
and simulated CSF flow using the Stokes equations.
We identified that the velocities were most strongly
affected when removing cilia in the dorsal diencephalic cilia,
as the maximum velocity dropped from 27.7 $\mu$m/s to 8.0 $\mu$m/s.
In contrast, removing telencephalic and ventral diencephalic had minimal
impact and resulted in maximum velocities of 27.7 $\mu$m/s or 27.2 $\mu$m/s, respectively.

With the resulting velocity fields, we simulated transport of Dendra2.
Changes in the transport dynamics were most prominent when removing the dorsal cilia
in the diencephalic ventricle, which produces the strongest flow at this developmental stage.
All ROIs (except ROI 1, where we locally released solutes) exhibit significant changes
in the dynamics of $\cbar$ (\Cref{fig:figure5_compare_cilia_modifications}b).
Compared to the baseline model results, we observed a slower increase and lower final concentration
in the middle and anterior ventricles (ROIs 2--5),
and a faster transport to the posterior ventricle (ROI 6).
In contrast, removal of telencephalic cilia did not alter the evolution of the
mean concentration $\cbar$ with time (\Cref{fig:figure5_compare_cilia_modifications}b).
Finally, without ventral cilia in the diencephalic ventricle,
more solute has spread towards the anterior ventricle (ROI 5) and
less solute has spread towards the posterior ventricle (ROI 6).
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure5_compare_cilia_modifications.png}
    \caption{\textbf{a.} The three separate ciliated regions where loss of cilia motility is simulated
    by imposing a zero tangential traction boundary condition in the respective region.
    \textbf{b.} Mean concentration $\cbar(t)$ for the six regions of interest in the three 
    loss-of-cilia-motility scenarios considered. The region named in the legend is the region
    where we modify the cilia tangential traction boundary condition to a free-slip condition
    in the CSF flow model.}
    \label{fig:figure5_compare_cilia_modifications}
\end{figure}
